### 2023-05-02 Thèse/Autres/Abstract.tex
Mise en forme globale

\thispagestyle{empty}

\noindent \textbf{Abstract:}

\noindent TEXTE OF THE ABSTRACT

\vfill

\noindent \textbf{Résumé :}

\noindent TEXTE DU RESUME