### 2023-05-02 Thèse/Autres/Pagedegarde.tex

\thispagestyle{empty}

\newgeometry{left=2cm,right=2cm,top=3cm,bottom=3cm}

\includepdf[pages=-]{Autres/PDG.pdf}